\documentclass[20pt,landscape,pdftex]{foils}

\usepackage{ppftlk}
\renewcommand{\P}{\text{P}}
\newcommand{\MC}{\multicolumn}
\renewcommand\r{\right}
\renewcommand\l{\left}
\newcommand\dist{\buildrel\rm d\over\sim}
\newcommand\ind{\stackrel{\rm indep.}{\sim}}
\newcommand{\cY}{{\cal Y}}
\newcommand{\cT}{{\cal T}}
\setlength{\doublerulesep}{\arrayrulewidth}
\newtheorem{assumption}{Assumption}
\newcommand{\indep}{{\bot\negthickspace\negthickspace\bot}} 
\newcommand\spacingset[1]{\renewcommand{\baselinestretch}%
{#1}\small\normalsize}

%
\title{Matching as Nonparametric Preprocessing for Improving
  Parametric Causal Inference}

\date{August, 2004}
\author{Daniel E. Ho \\
  Yale Law School\\
  \\
  Kosuke Imai\\
  Department of Politics\\ Princeton University \\
  \\ 
  Gary King\\
  Department of Government\\ Harvard University\\
  \\
  Elizabeth A. Stuart\\
  Department of Statistics\\ Harvard University
\mbox{}\pdfbookmark{TitlePage}{stlab0}}

\parindent=0pt
\begin{document}
\color{black}
\LOGOOFF
%\hypersetup{pdfpagetransition={Split /Dm /H /M /O}}
%\hypersetup{pdfpagetransition={Blinds /Dm /H}}
\hypersetup{pdfpagetransition={Box /M /O}}
%\hypersetup{pdfpagetransition==Dissolve}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\foilhead{Introduction\pause}

\hypersetup{pdfpagetransition=Replace}

\begin{itemize}
\item Matching, a new statistics literature on causal inferences:\pause
  \begin{enumerate}
  \item nonparametric, non-model based methods.\pause
  \item reducing or eliminating functional form and other assumptions.\pause
  \end{enumerate}
  
\item But, from the point of view of practical researchers,\pause
  \begin{enumerate}    
  \item conflicting techniques, practices, guidelines, and rules of thumbs. \pause
  \item calculation of valid standard errors is complicated or
    unavailable.\pause
  \item few theoretical results.\pause
  \end{enumerate}

\item Our unifying idea of this diverse literature:\pause 
  \begin{enumerate}
  \item use matching NOT as a substitute for familiar parametric
    regression analysis.\pause
  \item use matching to make our parametric models work better.\pause
  \end{enumerate}

\item Under our proposed framework,\pause
  \begin{enumerate}
  \item parametric analyses are applied to preprocessed data rather
    than to raw data.\pause
  \item valid standard errors are readily available.\pause
  \item resulting estimates are less model dependent.\pause
  \end{enumerate} 

\item Easy-to-use software: MatchIt available at http://gking.harvard.edu/matchit/
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\foilhead{Definition of Causal Effects\pause}

\hypersetup{pdfpagetransition=Replace}


\end{document}