
\section{What \MatchIt\ Does}

\MatchIt\ implements the suggestions of \citet*{HoImaKin05} for
improving parametric statistical models and reducing model dependence
by preprocessing data with semi-parametric and non-parametric matching
methods.  After preprocessing with \MatchIt, researchers can use
whatever parametric model and software they would have used without
\MatchIt, without other modification, and produce inferences that are
substantially more robust and less sensitive to modeling assumptions.
(In addition, you may wish to use
\hlink{Zelig}{http://gking.harvard.edu/zelig/} \citep{ImaKinLau04} for
subsequent parametric analyses, as it is designed for maximum
convenience in analyzing \MatchIt\ datasets.)  \MatchIt\ reduces the
dependence of causal inferences on commonly made, but hard-to-justify,
statistical modeling assumptions via a wide range of sophisticated
matching methods.  In addition, we have written \MatchIt\ so that
adding new matching methods to the software is as easy for anyone with
the inclination as it is for us.

\section{Software Requirements} 
\label{sec:require}

\MatchIt\ works in conjunction with the R programming language and
statistical software, and will run on any platform where R is
installed (Windows, Unix, or Mac OS X).  R is available free for
download at the Comprehensive R Archive Network (CRAN) at
\hlink{http://cran.r-project.org/}{http://cran.r-project.org/}.
\MatchIt\ has been tested on the most recent version of R.  A good way
to learn R, if you don't know it already, is to learn Zelig (available
at \hlink{http://gking.harvard.edu}{http://gking.harvard.edu}) which
includes a self-contained introduction to R and can be used to analyze
the matched data after running \MatchIt.

\section{Installing \MatchIt}
\label{sec:install}

To install \MatchIt\ for all platforms, type at the R command prompt,
\begin{verbatim}
> install.packages("MatchIt")
\end{verbatim}
and \MatchIt\ will install itself onto your system automatically.
During the installation process you may either decide to keep or
discard the installation files, which will not affect the way
\MatchIt\ runs.  Note that you only need to do this once.  
%Some users
%may also find it helpful to install the package with version control
%(see Subsection~\ref{subsec:vercontrol}). 
We recommend that users also install Zelig, Matching, and optmatch by
typing the following commands,
\begin{verbatim}
> install.packages("Zelig")
> install.packages("Matching")
> install.packages("optmatch", contriburl = "http://www.stat.lsa.umich.edu/~bbh/optmatch")
\end{verbatim}


\section{Loading \MatchIt} \label{sec:load}

As with any R package, you need install \MatchIt\ only once, but you
must load it prior to each use.  You can do this for each R session by
typing,
\begin{verbatim}
> library(MatchIt)
\end{verbatim}
at the R command prompt.  

Alternatively, you can specify R to load \MatchIt\ automatically at
launch so that you can skip the step of typing {\tt library(MatchIt)}
at the beginning of every R session.  To do this, edit the {\tt
  Rprofile} file located in the R program subdirectory, e.g.
\texttt{C:/R/rw2011/etc/}, for Windows systems or the {\tt .Rprofile}
file located in the home directory for Unix/Linux and Mac OS X
systems.  Using a text editor such as Windows notepad and emacs, add
the following line to the file,
\begin{verbatim}
options(defaultPackages = c(getOption("defaultPackages"), "MatchIt"))
\end{verbatim}
For this change to take effect, you need to restart R.

\section{Updating \MatchIt}

We recommend that you periodically update \MatchIt\ at the R prompt by typing:
\begin{verbatim}
> update.packages()
> library(MatchIt)
\end{verbatim}
which will update all the libraries including \MatchIt\ and load the
new version of \MatchIt.


%%% Local Variables: 
%%% mode: pdflatex
%%% TeX-master: "matchit"
%%% End: 
